### 2020-05-25 writeup/intro.tex

In this work, we consider multiple high-dimensional linear regression tasks as input and focus on predicting a particular task that only has limited amount of labeled data.
Concretely, each task consists of $n_i$ datapoints in space $\real^p$ organized as matrix $X_i\in\real^{n_i\times p}$, for $1\le i\le p$.
The labels of $X_i$ are given by $Y_i = X_i\beta_i + \varepsilon_i$, where $\beta_i$ denotes the ground truth parameters for task $i$ and $\varepsilon_i$ denotes i.i.d. random noise with mean zero and variance $\sigma^2$.
Importantly, we assume that the number of datapoints $n_i$ is a small constant $\rho_i$ times $p$ and that $p$ is large. \todo{rational}
%We shall assume that each task data follows a linear model, i.e. $y_i = X_i \beta_i + \varepsilon_i$, $1\le i\le k$.
%Here $\beta_i\in\real^p$ is the model parameter for the $i$-th task.
%Each row of $X_i\in\real^{n_i\times p}$ is assumed to be drawn i.i.d. from a fixed
%distribution with covariance matrix $\Sigma_i$.
We use a hard parameter sharing architecture that contains a shared body $B\in\real^{p\times r}$ for all tasks and a separate prediction head $\set{W_i \in \real^{r}}_{i=1}^k$ for each task \cite{R17,MTDNN19}.
%    \paragraph{Different covariates:}
This corresponds to minimizing the following optimization objective.
\begin{align}
	\label{eq_mtl}
	f(B; W_1, \dots, W_k) = \sum_{i=1}^k \norm{X_i B W_i - Y_i}^2.
\end{align}
%Note that we consider the natural parameterization without reweighting the tasks above.
%The shared body $B$ plays an important role because it allows information transfer between different task data.
%This is known as the hard parameter sharing architecture in the literature, where we control the capacity $r$ of $B$, e.g. \cite{KD12,WZR20}.
%We focus on comparing the test performance on a particular task solved with equation \eqref{eq_mtl} to single-task learning.
%The details are described in Algorithm \ref{alg_estimator}.
Let $\hat{\beta}_t^{\MTL} = B W_t$ denote the optimal predictor obtained from solving equation \eqref{eq_mtl} for task $t$.
Our goal is to compare the test error of $\hat{\beta}_t^{\MTL}$, denoted as $\te(\hat{\beta}_t^{\MTL})$, to $\te(\hat{\beta}_t^{\STL})$, where $\hat{\beta}_t^{\STL}$ is likewise obtained from equation \eqref{eq_mtl} but with task $t$ in isolation.

>>>>

In this work, we consider multiple high-dimensional linear regression tasks as input and focus on predicting a particular task that only has limited amount of labeled data.
Concretely, each task consists of $n_i$ datapoints in space $\real^p$ organized as matrix $X_i\in\real^{n_i\times p}$, for $1\le i\le t$.
The labels of $X_i$ are given by $Y_i = X_i\beta_i + \varepsilon_i$, where $\beta_i$ denotes the ground truth parameters for task $i$ and $\varepsilon_i$ denotes i.i.d. random noise with mean zero and variance $\sigma^2$.
Importantly, we assume that the number of datapoints $n_i$ is a small constant $\rho_i$ times $p$ and that $p$ is large. \todo{rational}
%We shall assume that each task data follows a linear model, i.e. $y_i = X_i \beta_i + \varepsilon_i$, $1\le i\le k$.
%Here $\beta_i\in\real^p$ is the model parameter for the $i$-th task.
%Each row of $X_i\in\real^{n_i\times p}$ is assumed to be drawn i.i.d. from a fixed
%distribution with covariance matrix $\Sigma_i$.
We use a hard parameter sharing architecture that contains a shared body $B\in\real^{p\times r}$ for all tasks and a separate prediction head $\set{W_i \in \real^{r}}_{i=1}^t$ for each task \cite{R17,MTDNN19}.
%    \paragraph{Different covariates:}
This corresponds to minimizing the following optimization objective.
\begin{align}
	\label{eq_mtl}
	f(B; W_1, \dots, W_t) = \sum_{i=1}^t \norm{X_i B W_i - Y_i}^2.
\end{align}
%Note that we consider the natural parameterization without reweighting the tasks above.
%The shared body $B$ plays an important role because it allows information transfer between different task data.
%This is known as the hard parameter sharing architecture in the literature, where we control the capacity $r$ of $B$, e.g. \cite{KD12,WZR20}.
%We focus on comparing the test performance on a particular task solved with equation \eqref{eq_mtl} to single-task learning.
%The details are described in Algorithm \ref{alg_estimator}.
Let $\hat{\beta}_t^{\MTL} = B W_t$ denote the optimal predictor obtained from solving equation \eqref{eq_mtl} for task $t$.
Our goal is to compare the test error of $\hat{\beta}_t^{\MTL}$, denoted as $\te(\hat{\beta}_t^{\MTL})$, to $\te(\hat{\beta}_t^{\STL})$, where $\hat{\beta}_t^{\STL}$ is likewise obtained from equation \eqref{eq_mtl} but with task $t$ in isolation.

### 2020-05-25 writeup/setup.tex

Let $\hat{\beta}_t^{\STL} = (X_t^{\top}X_t)^{-1}X_t^{\top}Y_t$ denote the linear regression estimator, which can be also be obtained from equation \eqref{eq_mtl} by using the target task alone.
The bias-variance trade-off \cite{HTF09} says that
	\[ \te_t(\hat{\beta}) =
		\bignorm{\exarg{\varepsilon_t}{\hat{\beta}} - \beta_t}^2 + \exarg{\varepsilon_t}{\hat{\beta} - \exarg{\varepsilon_t}{\hat{\beta}}}. \]
Recall that we assume $n_i > p$ for every $1\le i\le t$.
Hence the bias of $\te_t(\hat{\beta}_t^{\STL})$ is zero and its test error is equal to variance, given by $\sigma^2 \cdot \tr{(X_i^{\top}X_i)^{-1}\Sigma_i}$.

For the case of two tasks, we decompose the test error of $\hat{\beta}_{t}^{\MTL}$ on the targe task into two parts
\begin{align}
	\te(\hat{\beta}_t^{\MTL}) =& \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_s - \hat{v}\beta_t)}^2 \nonumber \\
	&+ \sigma^2\cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}, \label{eq_te_model_shift}
\end{align}
where $\hat{v}$ denotes the ratio of the output layer weights (to be defined more precisely in Appendix \ref{app_proof_main}).

For the case of two tasks, we can get precise rates using random matrix theory.
For the sake of clarity, we call task 1 the source task and task 2 the target task,
i.e. $\beta_1 = \beta_s$ and $\beta_2 = \beta_t$.
We introduce the following notations for the high-dimensional setting
\[ c_{n_1} \define \frac{n_1}{p} \to c_1, \quad c_{n_2} \define \frac{n_2}p \to c_2, \quad \text{as } \ n_1, n_2\to \infty, \]
for some constants $c_1, c_2 \in (1,\infty)$.
A crucial quantity is what we call the \textit{covariate shift} matrix $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$.
Let $\lambda_1, \lambda_2, \dots, \lambda_p$ denote the singular values of $M$.
%

\textbf{Model shift bias.}

We begin by observing that the test error of $\hat{\beta}_t^{\MTL}$ consists of two parts.
One part captures how similar the task models are and the other part captures the variance of $\hat{\beta}_t^{\MTL}$.
Compared with $\hat{\beta}_t^{\STL}$, we observe that the variance part of $\hat{\beta}_t^{\MTL}$ gets reduced, since more data is added from source tasks.
The bias part of $\hat{\beta}_t^{\MTL}$, which we term as \textit{model shift bias}, affects performance negatively.
We derive the asympotic limit of $\te(\hat{\beta}_t^{\MTL})$ as $p$ approaches infinity.
We compare it with the asympotic limit of $\te(\hat{\beta}_t^{\STL})$, for settings where the target data size is limited.
We show sharp generalization bounds for two settings: i) two tasks with general covaraites; ii) many tasks with the same covariates.

\textbf{Variance reduction.}

It is not hard to see that the variance of $\hat{\beta}_t^{\MTL}$ is reduced compared to $\hat{\beta}_t^{\STL}$, i.e. %(following the argument of Proposition \ref{prop_monotone}), i.e.
\[ \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} \le \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}. \]
Because of model shift bias, we can no longer guarantee that $\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL})$.

>>>>

Let $\hat{\beta}_t^{\STL} = (X_t^{\top}X_t)^{-1}X_t^{\top}Y_t$ denote the linear regression estimator, which can be also be obtained from equation \eqref{eq_mtl} by using the target task alone.
The bias-variance trade-off \cite{HTF09} says that
	\[ \te_t(\hat{\beta}) =
		\bignorm{\exarg{\varepsilon_t}{\hat{\beta}} - \beta_t}^2 + \exarg{\varepsilon_t}{\hat{\beta} - \exarg{\varepsilon_t}{\hat{\beta}}}. \]
Recall that we assume $n_i > p$ for every $1\le i\le t$.
Hence the bias of $\te_t(\hat{\beta}_t^{\STL})$ is zero and its test error is equal to variance, given by $\sigma^2 \cdot \tr[(X_i^{\top}X_i)^{-1}\Sigma_i]$.
We focus on a setting where the data size $n_t$ of the target task is equal to a small constant $\rho_t > 1$ times $p$.
This setting captures the need for adding more labeled data to reduce the test error of the target task.

To illustrate our intuition, we begin by considering the multi-task learning setting with one source task.
Recall that $\hat{\beta}_t^{\MTL}$ is defined as $BW_t$ after solving equation \eqref{eq_mtl}.
We decompose the test error of $\hat{\beta}_{t}^{\MTL}$ on the targe task into two parts (to be derived in Appendix \ref{app_proof_main})
\begin{align}
	\te_t(\hat{\beta}_t^{\MTL}) =& ~ \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_1 - \hat{v}\beta_2)}^2 \label{eq_te_model_shift} \\
	&+ \sigma^2\cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}, \label{eq_te_var}
\end{align}
where $\hat{v} = W_2 / W_1$ denotes the ratio of the output layer weights.
By comparing the bias and variance of $\te_t(\hat{\beta}_t^{\MTL})$ to $\te_t(\hat{\beta}_t^{\STL})$, we observe the following two effects, respectively.

\textbf{Model-shift bias.}
%We begin by observing that the test error of $\hat{\beta}_t^{\MTL}$ consists of two parts.
We term equation \eqref{eq_te_model_shift} as \textit{model-shift bias}, which captures how similar $\beta_1$ and $\beta_2$ are.
This part introduces a negative effect to multi-task learning.
The scaling term $\hat{v}$ corresponds to the intuition that the shared module $B$ learns a subspace for both tasks.
The output layer of each task scales the direction of $B$ suitably to fit the task.

\textbf{Variance reduction.}
Next, it is not hard to verify that equation \eqref{eq_te_var}, the variance of $\hat{\beta}_t^{\MTL}$, is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
This part introduces a positive variance reduction effect to performing multi-task learning.
Hence, whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the trade-off between two effects:
(i) the positive effect from variance reduction;
(ii) the negative effect from model shift bias.

%, i.e. %(following the argument of Proposition \ref{prop_monotone}), i.e.
%\[ \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} \le \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}. \]
%Because of model shift bias, we can no longer guarantee that $\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL})$.

\textbf{Notations.} For the sake of clarity, we denote the first task as the source task and last ($t$-th) task as the target task.
